---
title: "Activity Analysis"
author: "Sarathy Iyer"
date: "February 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

# Header 1 Assignment practice Reproducible Research using Knitr 
# Load Libraries required for processing

```{r activity}

library(nlme)
library(lattice)
library(ggplot2)
library(dplyr)
```
## Loading and preprocessing the data

```{r readdata}
 activityData<-read.csv("./data/ReportPublish1/activity.csv")
 head(activityData)
 activityData1 <- activityData[ with (activityData, { !(is.na(steps)) } ), ]
```

## Summary of the 2 data sets to see the difference
```{r summary}
summary(activityData)
summary(activityData1)
str(activityData)
str(activityData1)
```

## Activity 1.0: Make a histogram of the total number of steps taken each day
### Activity 1.1:Group the activity data by date and sum the steps. 
### Display the steps_by_day
### Draw the histogram 
```{r histogram1, echo=TRUE}
by_day <- group_by(activityData1, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day
hist(steps_by_day$total, main="Histogram - total number of steps per day", 
     xlab="Total number of steps in a day")
```	 
### Activity 1.2. Calculate and report the mean and median total number of steps taken per day

```{r MeanAndMedianFrom Summary}
summary(steps_by_day)
```

## Activity 2.0: What is the average daily activity pattern?
### Activity 2.1: Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r timeseries5mininterval, echo=TRUE}
	avgsteps<-aggregate(steps ~ interval, data = activityData1, mean)
	plot(avgsteps$interval,avgsteps$steps,type="l",main="Average Steps in the interval",xlab="Interval",ylab="Average Steps")

```

### Activity 2.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r averageAndMaximumSteps}
  avgstepsdesc<-arrange(avgsteps,desc(steps))
	head(avgstepsdesc)
	maxinterval<-avgstepsdesc[1,1]
  maxinterval
```

## Activity 3.0 Imputing missing values
### Activity 3.1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r missingvalues}
xy1<-activityData[!complete.cases(activityData),]
	nrow(xy1)
```

### Activity 3.2 Devise a strategy for filling in all of the missing values in the dataset.
```{r FillMissingValues}
activityDataNoNA<-activityData
	for(i in 1:nrow(activityDataNoNA)) {
	  if ( is.na(activityDataNoNA$steps[i]) ){
		activityDataNoNA$steps[i] = avgsteps[activityDataNoNA$interval[i] == avgsteps$interval,2]
	  }
	}
```

### Activity 3.3 Create a new dataset that is equal to the original dataset but with the missing data filled in. 
activityDataNoNA is the new data set and activityData is the original 

### Activity 3.4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r NewHistogramwithImputValues, echo=TRUE}
by_day_nona <- group_by(activityDataNoNA, date)
	   steps_by_day_nona <- summarise(by_day_nona, total = sum(steps))
	   steps_by_day_nona
		hist(steps_by_day_nona$total, main="Histogram - total number of steps per day", xlab="Total number of steps in a day")
		
```

### Activity 3.5 What is the impact of imputing missing data on the estimates of the total daily number of steps?
Not much of difference in mean and median there is very mild difference of one
		
Summary of Imput Value 
```{r summariesForComparison}
summary(steps_by_day_nona)
summary(steps_by_day)
```

Here is the comparison the totals	with mean input vs original	`r steps_by_day_nona$total`
vs 
`r steps_by_day$total`
more number of totals are displayed. But it did not affect the histogram by viewing it


### Activity 4.0 Are there differences in activity patterns between weekdays and weekends?
### Activity 4.1 For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

```{r Weekday}
#weekdays(as.Date(activityDataNoNA$date,"%Y-%m-%d"))
```


### Activity 4.2
```{r facaddition}
for(i in 1:nrow(activityDataNoNA)) {
	    x=weekdays(as.Date(activityDataNoNA$date[i],"%Y-%m-%d") )
	  if ( x == "Saturday" || x=="Sunday") {
		activityDataNoNA$weekDayType[i] = "weekend"
	  }
	  else {
	     activityDataNoNA$weekDayType[i] = "weekday"
	  }
	 }
```


### Activity 4.3 Make a panel plot containing a time series plot based on the weekday weekend and steps
	
```{r FinalPlot, echo=TRUE}
	avgstepsNew<-aggregate(steps ~ interval+weekDayType, data = activityDataNoNA, mean)
    xyplot(steps ~ interval|weekDayType,activityDataNoNA, type="l", aspect = "iso")
	  
```

